---
title: "Identity Encoding Code 10"
subtitle: "Mantel Tests"
author: |
  <hr>
  <center style="font-style:normal;">
  <a style="font-size:22px;color:#337ab7;text-decoration: underline;"href="http://smith-vidaurre.com/">Grace Smith-Vidaurre</a><sup><span style="font-size:12px;color:black;text-decoration:none!important;">1-4*</span></sup>
  <br>
  <br>
  <center style="font-size:18px;font-style:normal;color:black;"><sup><span style="font-size:12px;color:black;">1</span></sup>Department of Biology, New Mexico State University</center>
  <center style="font-size:18px;font-style:normal;color:black;"><sup><span style="font-size:12px;color:black;">2</span></sup>Laboratory of Neurogenetics of Language, Rockefeller University</center>
  <center style="font-size:18px;font-style:normal;color:black;"><sup><span style="font-size:12px;color:black;">3</span></sup>Field Research Center, Rockefeller University</center>
  <center style="font-size:18px;font-style:normal;color:black;"><sup><span style="font-size:12px;color:black;">4</span></sup>Department of Biological Sciences, University of Cincinnati</center>
  <br>
  <center style="font-size:18px;"><sup style="font-size:12px;">*</sup>gsvidaurre@gmail.com</center>
  <br>
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: false
---

<style type="text/css">

a:hover {
  color: #23527c !important;
}

h1.title { /* Document title */
  font-size: 32px;
  color: black;
  font-weight: normal;
  text-align: center;
}

h1 {
   color: #0E0E7D;
   font-size: 26px;
   font-weight: normal;
}

h2 {
   color: #0E0E7D;
   font-size: 24px;
   font-weight: bold;
}

h3 { /* Document subtitle */
   color: #0E0E7D;
   font-size: 28px;
   font-weight: normal;
   text-align: center;
}

h4 {
   color: #0E0E7D;
   font-size: 20px;
   font-weight: normal;
}

h4.date { /* Date in document header */
  font-size: 22px;
  font-style:normal;
  text-align: center;
}

body{ /* Normal */
      font-size: 20px;
  }
  
code.r{ /* Code block */
    font-size: 20px;
}
</style>

```{r global options, include = FALSE}

knitr::opts_knit$set(root.dir = "/home/gsvidaurre/Desktop/GitHub_repos/identity-information-post-introduction", echo = TRUE, include = TRUE, eval = TRUE)

knitr::opts_chunk$set(root.dir = "/home/gsvidaurre/Desktop/GitHub_repos/identity-information-post-introduction", echo = TRUE, include = TRUE, eval = TRUE)

```

<hr>

This code accompanies the following paper in PLOS Computational Biology:

Smith-Vidaurre, G., Perez-Marrufo, V., Hobson, E.A., Salinas-Melgoza, A., and T.F. Wright. Individual identity information persists in learned calls of introduced parrot populations.

<hr>

Mantel tests were conducted per range to evaluate the statistical significance of acoustic convergence across social scales. For the site scale, Mantel tests were conducted over similarity methods: spectrographic cross-correlation (SPCC) and random forests, as well as over the 3 datasets used to address potential repeated sampling of individuals: no filtering (full dataset), filtering by Gaussian mixture models, and filtering by visual classification. See the associated publication for more information about these analyses.

Check out Github repositories from previous work for related analyses and code:

- [gsvidaurre/strong-individual-signatures](https://github.com/gsvidaurre/strong-individual-signatures)<a href='#References'><sup>[1]</sup></a>

- [gsvidaurre/simpler-signatures-post-invasion](https://github.com/gsvidaurre/simpler-signatures-post-invasion)<a href='#References'><sup>[2]</sup></a>

Please cite the associated papers as well as our code (see DOIs on GitHub) if the code or analyses across these repositories are useful for your own research.
```{r message = FALSE}

rm(list = ls())

X <- c("tidyverse", "pbapply", "plyr", "dplyr", "data.table", "parallel", "vegan", "knitr")

invisible(lapply(X, library, character.only = TRUE))

# Suppress summarise warnings from dplyr
options(dplyr.summarise.inform = FALSE)

path <- "/media/gsvidaurre/MYIOPSITTA/R/VocalLearning_PostDisruption/Data"
gpath <- "/media/gsvidaurre/MYIOPSITTA/R/VocalLearning_PostDisruption/Graphics"
seed <- 401
cores <- parallel::detectCores() - 2

```

Read in the extended selection table (EST) that contains metadata and wave objects for pre-processed native and introduced range calls across the individual and site scales.
```{r}

nat_int_est <- readRDS(file.path(path, "monk_parakeet_contactCalls_rangeComparison_extSelTable.RDS"))
glimpse(nat_int_est)

```

# Filtering out some calls

I dropped individual scale calls that had been added to the site scale (1 call per known repeatedly sampled individual in the final individual scale dataset, suffix "site_scale").
```{r chunk3}

# 31 calls with the suffix "_site_scale". Not all of these belonged to the individuals in the final individual scale dataset, since some individuals that we repeatedly sampled were dropped due to low sample sizes after pre-processing
nat_int_est %>%
  dplyr::filter(social_scale == "Site") %>% 
  dplyr::filter(grepl("_site_scale", sound.files)) %>%
  pull(sound.files) %>%
  length()

# Get the "_site_scale" calls for the final dataset of known repeatedly sampled individuals
indiv_ids <- nat_int_est %>%
  dplyr::filter(social_scale == "Individual") %>%
  pull(Bird_ID) %>%
  unique()

# 17 total: 8 native and 9 introduced range
indiv_ids

# INT-UM1, INT-UM6, and INT-UM19 were each the sole bird sampled at sites BART, ASCA, and CAME, respectively. These sites were not included at the site scale, so these individuals did not have any "_site_scale" calls 
# This leaves 14 calls to drop, each representing a call for a different repeatedly sampled individual included in the site scale dataset
drop_is_calls <- nat_int_est %>%
  dplyr::filter(social_scale == "Site") %>%
  dplyr::filter(grepl("_site_scale", sound.files)) %>%
  dplyr::filter(Bird_ID %in% indiv_ids) %>%
  pull(sound.files)

length(drop_is_calls)

# Checking, looks good
nat_int_est %>%
  as_tibble() %>% 
  dplyr::filter(sound.files %in% drop_is_calls) %>%
  dplyr::select(sound.files, Bird_ID) %>%
  kable(align = rep("c", nrow(.)))

# Drop these calls from the extended selection table and continue with analyses
# nat_int_est <- nat_int_est %>%
  # dplyr::filter(!sound.files %in% drop_is_calls)

# Don't use tidyverse since I want to filter both the data frame and the .wav objects in the EST
nat_int_est <- nat_int_est[-grep(paste(paste("^", drop_is_calls, "$", sep = ""), collapse = "|"), nat_int_est$sound.files), ]

# 1582 calls remain out of the original 1596 calls read in above
dim(nat_int_est)
class(nat_int_est)

```

# Individual scale

We used 5 native range and 5 introduced range individuals for Mantel tests at the individual scale. Although these native range individuals were sampled at a single site (site 1145 sampled in 2017), and the introduced range individuals sampled at 3 sites (city of Austin in 2019), a previous validation analysis found that these individuals represented similar patterns of acoustic variation<a href='#References'><sup>[2]</sup></a>. When individuals from larger geographic distancs were included, calls overlapped more in acoustic space<a href='#References'><sup>[2]</sup></a>. Aside from the introduced range individuals recorded in Austin, most introduced range individuals were recorded over very large geographic distances. As such, we used native range birds from site 1145 and introduced range birds from Austin 2019 for this comparison of individual signatures in our dataset. 

Number of calls per native range individual at site 1145.
```{r echo = TRUE, eval = TRUE}

nat_int_est %>%
  dplyr::filter(social_scale == "Individual") %>%
  dplyr::filter(range == "Native") %>%
  dplyr::filter(site == "1145") %>%
  group_by(Bird_ID) %>%
  dplyr::summarise(n_calls = length(sound.files))

```

Number of calls per native range individual recorded in Austin 2019. 
```{r echo = TRUE, eval = TRUE}

nat_int_est %>%
  dplyr::filter(social_scale == "Individual") %>%
  dplyr::filter(range == "Introduced") %>%
  dplyr::filter(introduced_city == "Austin" & year == 2019) %>%
  group_by(Bird_ID) %>%
  dplyr::summarise(n_calls = length(sound.files))

```

# Mantel tests 

# Individual scale, SPCC 

This analysis addresses the first half of the hierachical mapping analysis (individual scale, site scale needs filtering and analyses run in triplicate for the potential repeated sampling issue). Mantel tests were performed with the individuals recorded over restricted geographic areas in each range (a single site-year for the native range, a single city-year for the introduced range). A partial Mantel test was conducted for the introduced range birds to control for site identity, as the introduced range birds were recorded at 3 different sites. We used only SPCC similarity for Mantel tests at the individual scale, as random forests models had been trained on individual scale calls.
```{r eval = FALSE}

xc_mat_spec <- readRDS(file.path(path, "xc_mat_nat_int_1582calls.RDS"))
dim(xc_mat_spec) == nrow(nat_int_est)
glimpse(xc_mat_spec)

# Iterate over ranges
range <- c("Native", "Introduced")
sts <- c("1145_2017", "SOCC_2019|ELEM_2019|INTR_2019")
perms <- 9999
file_nm <- "Mantel_results_indiv_scale.txt"

# A txt file will be populated by appending Mantel test results
file.remove(file.path(path, file_nm)) # remove previous versions as needed

# i <- 1 # testing
invisible(pblapply(1:length(range), function(i){
  
    # Subset the EST by the given range
    rng <- range[i]
    
    tmp_df <- nat_int_est %>%
      dplyr::filter(social_scale == "Individual") %>%
      dplyr::filter(range == rng) %>%
      dplyr::filter(grepl(sts[i], site_year)) %>%
      droplevels()
    # glimpse(tmp_df)
  
    # Get calls for the given iteration
    calls <- tmp_df %>%
      pull(sound.files)
    # head(calls)
  
    # Subset the SPCC matrix, convert to distance matrix
    xc_mat_tmp <- xc_mat_spec[grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(xc_mat_spec)[[1]]), grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(xc_mat_spec)[[2]])]
    # str(xc_mat_tmp)
  
    xc_mat_dist <- stats::as.dist(1 - xc_mat_tmp, diag = TRUE, upper = TRUE)
    # str(xc_mat_dist)
  
    # Make a binary identity matrix for individual identity
    # Predictor matrix of individual identity membership for Mantel tests
    id_indiv <- unlist(lapply(1:nrow(tmp_df), function(x){
        return(as.numeric(tmp_df$Bird_ID[x] == tmp_df$Bird_ID))
    }))

    id_mat <- matrix(id_indiv, nrow = nrow(tmp_df), ncol = nrow(tmp_df), dimnames = list(tmp_df$Bird_ID, tmp_df$Bird_ID))
    # str(id_mat)

    id_mat_dist <- stats::as.dist(1 - id_mat, diag = TRUE, upper = TRUE)
    # str(id_mat_dist)
  
    # Make a binary identity matrix for site-year identity if on the introduced range
    if(range[i] == "Introduced"){
      # Predictor matrix of site membership for partial Mantel tests
      id_site <- unlist(lapply(1:nrow(tmp_df), function(x){
          return(as.numeric(tmp_df$site_year[x] == tmp_df$site_year))
      }))

      part_mat <- matrix(id_site, nrow = nrow(tmp_df), ncol = nrow(tmp_df), dimnames = list(tmp_df$site_year, tmp_df$site_year))

      part_mat_dist <- stats::as.dist(1 - part_mat, diag = TRUE, upper = TRUE)
      # str(part_mat_dist)
    }
  
    # Perform a partial Mantel test for the introduced range only
    if(range[i] == "Native"){
      mntl <- vegan::mantel(xdis = id_mat_dist, ydis = xc_mat_dist, method = "pearson", permutations = perms, parallel = cores)
      partialM <- "N"
    } else if(range[i] == "Introduced"){
      mntl <- vegan::mantel.partial(xdis = id_mat_dist, ydis = xc_mat_dist, zdis = part_mat_dist, method = "pearson", permutations = perms, parallel = cores)
      partialM <- "Y"
    }
  
    # Make a data frame summarizing the output
    res_df <- data.frame(sim_method = "SPCC", range = range[i], num_indivs = length(unique(tmp_df$Bird_ID)), num_calls = dim(id_mat)[1], num_site_year = length(unique(tmp_df$site_year)), partial_Mantel = partialM, mantel_r = round(mntl$statistic, digits = 2), mantel_p = mntl$signif, permutations = perms)
  
    # If on the first iteration, write a new file
    if(i == 1){
      write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = TRUE, quote = FALSE, row.names = FALSE)

    # Append to the existing file in subsequent iterations
    } else {
      write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE, append = TRUE)
    }
  
    return(res_df)
  
}))

```

## Site scale

### General approach

We performed analyses to assess hierarchical mapping at the site scale while accounting for potential repeated sampling of individuals. Mantel tests were performed on 3 versions of the site scale data: 1) all calls per site-year, 2) calls filtered by clustering (Gaussian mixture modeling) results, and 3) calls filtered by multi-observer visual classification results. 

Mantel tests were performed to compare site scale convergence between ranges. For these tests, when a site or city was sampled in more than one year, we used only calls from the most recent sampling year. Calls recorded across years in the same introduced range site or city were used in Mantel tests over time (Austin and New Orleans only).

## Mantel tests per range

We performed Mantel tests for the three site-scale datasets and both similarity methods. We retained calls recorded in the cities of Austin and New Orleans in more recent sampling years. Austin and New Orleans calls from earlier sampling years were used in Mantel tests to compare site scale convergence over time (see below). I ran Mantel tests using SPCC and random forests similarity, as well as the 3 datasets generated above for the site scale. I also ran a Mantel test for the native range (no filtering for repeated individuals) as reported in our previous work<a href='#References'><sup>[1]</sup></a> with the 4 calls for UM4 retained at the site scale.

Read in similarity matrices and change dimension names for grepping below per Mantel test.
```{r}

# Read the SPCC similarity matrix for the full dataset of calls
xc_mat_spec <- readRDS(file.path(path, "xc_mat_nat_int_1582calls.RDS"))

# Read in random forests similarity (proximity matrix) for the validation and prediction call datasets
vp_prox_mat <- readRDS(file.path(path, "rf_validatn_predictn_prox_mat.RDS"))

```

Read in the 3 site scale datasets for this comparison between ranges.
```{r}

site_scale_nf <- readRDS(file.path(path, "site_scale_nf.RDS"))
site_scale_cf <- readRDS(file.path(path, "site_scale_cf.RDS"))
site_scale_vf <- readRDS(file.path(path, "site_scale_vf.RDS"))

```

Perform the Mantel tests by iterating over ranges, similarity methods, and data subsets representing approaches for addressing potential repeated sampling of individuals.
```{r eval = FALSE}

ranges <- c("Native", "Introduced")
sim_meth <- c("SPCC", "Random forests")
data_subset <- c("Full", "Clustering", "Visual inspection")

mat_list <- list(xc_mat_spec, vp_prox_mat)
df_list <- list(site_scale_nf, site_scale_vf, site_scale_cf)

perms <- 9999
file_nm <- "Mantel_results_site_scale_spatial.txt"

# A txt file will be populated by appending Mantel test results
file.remove(file.path(path, file_nm)) # remove previous versions as needed

# i <- 1
# j <- 1
# z <- 1 # testing
invisible(pblapply(1:length(ranges), function(i){
  
  lapply(1:length(sim_meth), function(j){
    
    lapply(1:length(data_subset), function(z){
    
      # Initialize the data frame of calls for the given site scale dataset
      ss_df <- df_list[[z]]
      # glimpse(ss_df)
      
      # Filter the extended selection table by calls in the given site scale dataset
      dats <- nat_int_est %>%
        as_tibble() %>%
        dplyr::filter(sound.files %in% ss_df$sound.files)
        # glimpse(dats)
      
      # Subset the data frame by the given range
      rng <- ranges[i]
      
      tmp_df <- dats %>%
        dplyr::filter(range == rng) %>%
        droplevels()
  
      # unique(tmp_df$range)
      
      # Get calls for the given range
      calls <- tmp_df %>%
        pull(sound.files)
      
      # head(calls, 25)
      
      # Initialize the correct similarity matrix
      sim_mat <- mat_list[[j]]
      
      # Subset the similarity matrix by the calls for the given range in the given dataset
      sim_mat_tmp <- sim_mat[grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[1]]), grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[2]])]
  
      # Convert to distance matrix by subtracting from 1
      # Convert to a dist object for the Mantel test
      sim_mat_dist <- stats::as.dist(1 - sim_mat_tmp, diag = TRUE, upper = TRUE)
      # str(sim_mat_dist)
  
      # Make a binary identity matrix for site-year identity
      # Predictor matrix of site-year identity membership for Mantel tests
      id_site_year <- unlist(lapply(1:nrow(tmp_df), function(y){
        return(as.numeric(tmp_df$site_year[y] == tmp_df$site_year))
      }))

      # Convert the vector into a similarity matrix
      id_mat <- matrix(id_site_year, nrow = nrow(tmp_df), ncol = nrow(tmp_df), dimnames = list(tmp_df$site_year, tmp_df$site_year))
      # head(id_mat)

      # Convert to a binary distance matrix of site-year identity
      id_mat_dist <- stats::as.dist(1 - id_mat, diag = TRUE, upper = TRUE)
      # str(id_mat_dist)
  
      # Perform the Mantel test
      mntl <- vegan::mantel(xdis = id_mat_dist, ydis = sim_mat_dist, method = "pearson", permutations = perms, parallel = cores)
  
      # Make a data frame summarizing the output
      res_df <- data.frame(method = sim_meth[j], range = ranges[i], dataset = data_subset[z], num_calls = nrow(tmp_df), num_site_year = length(unique(tmp_df$site_year)), partial_Mantel = "N", mantel_r = round(mntl$statistic, digits = 2), mantel_p = mntl$signif, permutations = perms)
  
    # If on the first iteration, write a new file
    if(i == 1 & j == 1 & z == 1){
      write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = TRUE, quote = FALSE, row.names = FALSE)

    # Append to the existing file in subsequent iterations
    } else {
      write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE, append = TRUE)
    }
      
    })
    
  })
  
}))

```

# Results across social scales and ranges

Adjust alpha of 0.05 by a Bonferroni correction to account for multiple testing in this comparison between ranges.
```{r}

# Original alpha 
alpha <- 0.05

mntl_indiv_df <- read.table(file.path(path, "Mantel_results_indiv_scale.txt"), sep = ",", header = TRUE)
nrow(mntl_indiv_df) # 2 tests

mntl_sites_df <- read.table(file.path(path, "Mantel_results_site_scale_spatial.txt"), sep = ",", header = TRUE)
nrow(mntl_sites_df) # 12 tests

# Adjust alpha by a Bonferroni correction using the number of comparisons at the individual and site scales
adj_alpha <- round(alpha/(nrow(mntl_indiv_df) + nrow(mntl_sites_df)), digits = 4)
adj_alpha # 0.0036

# Individual scale results
mntl_indiv_df %>%
  dplyr::select(-c(permutations, partial_Mantel)) %>%
  kable()

# Site scale results
mntl_sites_df %>%
  dplyr::select(-c(permutations, partial_Mantel)) %>%
  kable()

```

Individual and site scale p-values were significant for both ranges at the adjusted alpha.

### Validation of retention of UM4 calls for native range site scale

I ran Mantel tests for the native range site scale dataset with and without the UM4 calls, using both SPCC and random forests similarity.
```{r}

# Get the calls that were inadvertently retained for NAT-UM4 at the site scale
dropUM4 <- nat_int_est %>%
  as_tibble() %>% 
  dplyr::filter(social_scale == "Site") %>%
  dplyr::filter(Bird_ID == "NAT-UM4" & !grepl("site_scale", sound.files)) %>%
  pull(sound.files)
dropUM4

# Site scale calls, full dataset with NAT-UM4 calls
# Note that there are 602 calls here (610 in the extended selection table prior to dropping calls earlier in the script) because I added back one call each for a few more repeatedly sampled individuals in this work compared to [1]
# Regardless, the results with these calls can be compared with the native range results for the native range full dataset to assess whether the 4 calls for UM4 change the Mantel test statistic
site_scale_nf_natUM4 <- nat_int_est %>%
  as_tibble() %>% 
  dplyr::filter(social_scale == "Site") %>%
  dplyr::filter(range == "Native") %>%
  droplevels()
dim(site_scale_nf_natUM4)

# Site scale calls, full dataset without NAT-UM4 calls (606 calls)
site_scale_nf_nat <- site_scale_nf %>%
  as_tibble() %>% 
  dplyr::filter(range == "Native") %>%
  # Make sure to filter by calls in the filtered extended selection table
  dplyr::filter(sound.files %in% nat_int_est$sound.files) %>% 
  droplevels()
dim(site_scale_nf_nat)

# Checking, looks good
site_scale_nf_natUM4 %>% 
  dplyr::filter(sound.files %in% dropUM4) %>% 
  pull(sound.files)

site_scale_nf_nat %>% 
  dplyr::filter(sound.files %in% dropUM4) %>% 
  pull(sound.files)

sim_meth <- c("SPCC", "RF")
data_subset <- c("Full with UM4 calls", "Full without UM4 calls")

mat_list <- list(xc_mat_spec, vp_prox_mat)
df_list <- list(site_scale_nf_natUM4, site_scale_nf_nat)

perms <- 9999
  
# j <- 1
# z <- 1 # testing
res_df <- rbindlist(pblapply(1:length(sim_meth), function(j){
    
  res_df2 <- rbindlist(lapply(1:length(data_subset), function(z){
    
    tmp_df <- df_list[[z]]
    # glimpse(tmp_df)
      
    # Get calls
    calls <- tmp_df %>%
      pull(sound.files)
    # head(calls, 25)
    # length(calls)
      
    # Initialize the correct similarity matrix
    sim_mat <- mat_list[[j]]
      
    # Subset the similarity matrix by the calls for the given range in the given dataset
    sim_mat_tmp <- sim_mat[grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[1]]), grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[2]])]
    # str(sim_mat_tmp)
    dim(sim_mat_tmp)
  
    # Convert to distance matrix by subtracting from 1
    # Convert to a dist object for the Mantel test
    sim_mat_dist <- stats::as.dist(1 - sim_mat_tmp, diag = TRUE, upper = TRUE)
    # str(sim_mat_dist)
  
    # Make a binary identity matrix for site-year identity
    # Predictor matrix of site-year identity membership for Mantel tests
    id_site_year <- unlist(lapply(1:nrow(tmp_df), function(y){
      return(as.numeric(tmp_df$site_year[y] == tmp_df$site_year))
    }))

    # Convert the vector into a similarity matrix
    id_mat <- matrix(id_site_year, nrow = nrow(tmp_df), ncol = nrow(tmp_df), dimnames = list(tmp_df$site_year, tmp_df$site_year))
    # head(id_mat)
    # str(id_mat)

    # Convert to a binary distance matrix of site-year identity
    id_mat_dist <- stats::as.dist(1 - id_mat, diag = TRUE, upper = TRUE)
    # str(id_mat_dist)
  
    # Perform the Mantel test
    mntl <- vegan::mantel(xdis = id_mat_dist, ydis = sim_mat_dist, method = "pearson", permutations = perms, parallel = cores)
  
    # Make a data frame summarizing the output
    res_df <- data.frame(method = sim_meth[j], range = "Native", dataset = data_subset[z], num_calls = nrow(tmp_df), num_site_year = length(unique(tmp_df$site_year)), partial_Mantel = "N", mantel_r = round(mntl$statistic, digits = 4), mantel_p = mntl$signif, permutations = perms)
    
    return(res_df)
      
    }))
    
 return(res_df2)
  
}))

res_df %>%
  dplyr::select(-c(permutations, partial_Mantel)) %>%
  kable()

```

Mantel test statistics were very similar with and without the UM4 calls for the native range out to 4 significant digits, and are the same when rounded to 2 digits (and are very similar to those reported in previous work<a href='#References'><sup>[1]</sup></a>). Therefore, the inclusion of these UM4 calls at the site scale in previous work did not influence the findings previously reported.

# Mantel tests over time in the introduced range

Read in the 3 site scale datasets for this comparison over time in the introduced range.
```{r}

site_scale_nf_temp <- readRDS(file.path(path, "site_scale_nf_temp.RDS"))
site_scale_cf_temp <- readRDS(file.path(path, "site_scale_cf_temp.RDS"))
site_scale_vf_temp <- readRDS(file.path(path, "site_scale_vf_temp.RDS"))

```

Run the Mantel tests over similarity methods, years, introduced cities, datasets, to ask how similar calls were within versus among sites for the given year and city.
```{r eval = FALSE}

sim_meth <- c("SPCC", "Random forests")
years <- c("2004", "2011", "2019")
city <- c("Austin", "New Orleans")
data_subset <- c("Full", "Clustering", "Visual inspection")

mat_list <- list(xc_mat_spec, vp_prox_mat)

df_list <- list(
  site_scale_nf_temp,
  site_scale_vf_temp, 
  site_scale_cf_temp
)

perms <- 9999
file_nm <- "Mantel_results_site_scale_temporal.txt"

# A txt file will be populated by appending Mantel test results
file.remove(file.path(path, file_nm)) # remove previous versions as needed
 
# j <- 1
# x <- 1
# i <- 1
# z <- 1 # testing
invisible(pblapply(1:length(sim_meth), function(j){
  
  lapply(1:length(years), function(x){
    
    lapply(1:length(city), function(i){
    
      lapply(1:length(data_subset), function(z){
    
        # Get the right dataset and subset by the given year and city
        tmp_df <- df_list[[z]] %>%
          dplyr::filter(year == years[x]) %>%
          dplyr::filter(introduced_city == city[i]) %>%
          droplevels()
        # glimpse(dats)
      
        # unique(tmp_df$introduced_city)
        # unique(tmp_df$site_year)
        
        # Excludes New Orleans 2019, for which we do not have recordings
        if(nrow(tmp_df) > 0){
          
          # Get calls for the given comparison
          calls <- tmp_df %>%
            pull(sound.files)
          # head(calls, 25)
      
          # Initialize the correct similarity matrix
          sim_mat <- mat_list[[j]]
      
          # Subset the similarity matrix by the calls for the given range in the given dataset
          sim_mat_tmp <- sim_mat[grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[1]]), grep(paste(paste("^", calls, "$", sep = ""), collapse = "|"), dimnames(sim_mat)[[2]])]
  
          # Convert to distance matrix by subtracting from 1
          # Convert to a dist object for the Mantel test
          sim_mat_dist <- stats::as.dist(1 - sim_mat_tmp, diag = TRUE, upper = TRUE)
          # str(sim_mat_dist)
  
          # Make a binary identity matrix for site identity
          id_site <- unlist(lapply(1:nrow(tmp_df), function(y){
            return(as.numeric(tmp_df$site[y] == tmp_df$site))
          }))

          # Convert the vector into a similarity matrix
          id_site_mat <- matrix(id_site, nrow = nrow(tmp_df), ncol = nrow(tmp_df), dimnames = list(tmp_df$site, tmp_df$site))
          # head(id_site_mat)

          # Convert to a binary distance matrix of site-year identity
          id_site_mat_dist <- stats::as.dist(1 - id_site_mat, diag = TRUE, upper = TRUE)
          # str(id_site_mat_dist)
  
          # Perform the partial Mantel test for the effect of year on call similarity while controlling for site identity
          mntl <- vegan::mantel(xdis = id_site_mat_dist, ydis = sim_mat_dist, method = "pearson", permutations = perms, parallel = cores)
  
          # Make a data frame summarizing the output
          res_df <- data.frame(method = sim_meth[j], year = years[x], city = city[i], dataset = data_subset[z], num_calls = nrow(tmp_df), num_sites = length(unique(tmp_df$site)), partial_Mantel = "N", mantel_r = round(mntl$statistic, digits = 2), mantel_p = mntl$signif, permutations = perms)
  
        # If on the first iteration, write a new file
        if(j == 1 & x == 1 & i ==1 & z == 1){
          write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = TRUE, quote = FALSE, row.names = FALSE)

        # Append to the existing file in subsequent iterations
        } else {
          write.table(res_df, file = file.path(path, file_nm), sep = ",", col.names = FALSE, quote = FALSE, row.names = FALSE, append = TRUE)
        }
          
        }
      
      })
    
    })
    
  })
  
}))

```

Adjust alpha of 0.05 by a Bonferroni correction to account for multiple tests over time.
```{r}

# Original alpha 
alpha <- 0.05

mntl_temp_df <- read.table(file.path(path, "Mantel_results_site_scale_temporal.txt"), sep = ",", header = TRUE)
nrow(mntl_temp_df) # 30 tests

# Adjust alpha by a Bonferroni correction using the number of comparisons (one per year)
adj_alpha <- round(alpha/nrow(mntl_temp_df), digits = 4)
adj_alpha # 0.0017

mntl_temp_df %>%
  dplyr::select(-c(permutations, partial_Mantel)) %>%
  # View()
  kable()

```

# References

    1. Smith-Vidaurre, G., Araya-Salas, M., and T.F. Wright. 2020. Individual signatures outweigh social group identity in contact calls of a communally nesting parrot. Behavioral Ecology 31(2), 448-458. https://doi.org/10.1093/beheco/arz202
    
    2. Smith-Vidaurre, G., Perez-Marrufo, V., & Wright, T. F. 2021. Individual vocal signatures show reduced complexity following invasion. Animal Behavior, 179, 15–39. https://doi.org/10.1016/j.anbehav.2021.06.020

Documenting session information and software versions at the time of knitting the RMarkdown output.
```{r}

sessionInfo()

```